---
title: "analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r, warning = FALSE}
library(tidyverse)
```
```{r}
confi = "../../configurations/configurations_membership_dataset.csv"
confi = read.csv(confi)
```


### accuracy

```{r}
accuracy = data.frame(matrix(nrow = 0, ncol = 7))

for (set_index in 1:81){
  for (set_type in c('balanced', 'unbalanced')){
    res_path = paste0('../res/', set_type, '_', set_index, '/')
    # read true Y
    Y = paste0(res_path, 'data/test_set/Y.csv') %>% read.csv(header = FALSE) 
    Ntest = dim(Y)[1]
    truek0 = confi$k0[set_index]
    truek1 = confi$k1[set_index]
    
    # learn with true topics
    if (file.exists(paste0(res_path, 'true/Y_', set_index, '.csv'))){
      temY = paste0(res_path, 'true/Y_', set_index, '.csv') %>% read.csv(header = FALSE)
      accu = sum(temY == Y) / Ntest
      accuracy = rbind(accuracy, data.frame(set_index, set_type, accu, setting = 'truth', k0 = truek0, k1 = truek1, method = 'supervised'))
    }
    if (file.exists(paste0(res_path, 'true/Y_un.csv'))){
      temY = paste0(res_path, 'true/Y_un.csv') %>% read.csv(header = FALSE) 
      accu = sum(temY == Y) / Ntest
      accuracy = rbind(accuracy, data.frame(set_index, set_type, accu, setting = 'truth', k0 = truek0, k1 = truek1, method = 'unsupervised'))
    }
    
    for (k0s in c(2, 5, 10)){
      for (k1s in c(2, 5, 10)){
        for (meth in c('supervised', 'unsupervised')){
          tem1 = paste0(res_path, meth, '_', k0s, '_', k1s)
          if (tem1 %>% dir.exists()){
            temY = paste0(tem1, '/Y.csv') %>% read.csv(header = FALSE)
            accu = sum(temY == Y) / Ntest
            accuracy = rbind(accuracy, data.frame(set_index, set_type, accu, setting = 'training', k0 = k0s, k1 = k1s, method = meth))
          }
        }
      }
    }    
  }
}
```

```{r}
accuracy = confi %>% 
  select(set_index, k0, k1, nlabel, d) %>%
  rename(truek0 = k0, truek1 = k1) %>%
  merge(accuracy, by = "set_index") 
```

```{r}
accuracy %>%
  filter(k0 == truek0 & k1 == truek1 & d == 50) %>%
  ggplot(aes(x = factor(set_index), y = accu, color = setting)) + geom_point() + facet_grid(method~set_type)
```

Observation: 
 - Overall, the test accuracy is closed to the accuracy when we know the true topics. That means our training process is correct and the low accuracy is not caused by computation.



```{r}
temp2 = accuracy %>%
  filter(setting == 'training') %>%
  filter(truek0 == truek1 & k0 == k1) %>%
  filter(truek0 == k0 & truek1 == k1)

for (d0 in c(20, 50, 100)){
  p = accuracy %>%
    filter(setting == 'training') %>%
    filter(truek0 == truek1 & k0 == k1) %>%
    filter(d == d0) %>%
    ggplot(aes(x = k0, y = accu, color = method, group = interaction(method,set_index))) + 
    geom_path() + 
    facet_grid(method ~ set_type) +
    geom_point(data = temp2 %>% filter(d == d0), aes(x = k0, y = accu, color = 'red')) + 
    ggtitle(paste0("d = ", d0))
  print(p)
}


```

Observations:
 - Supervised model is robust to misspecified topic numbers.
 - For unsupervised model, it's even better to train it with # topics smaller than the true parameters.
 - Overall, high dimension will improve the accuracy.



### membership


```{r}
# set 10:
# k0 = k1 = 2, d = 20, nlabel = 6
set_index = 10
set_type = 'balanced'
res_path = paste0('../res/', set_type, '_', set_index, '/')

accuracy %>% filter(set_index == 10 & set_type == 'balanced') %>% select(k0, k1, setting, method, accu) %>% ggplot(aes(x = k0, y = accu, color = method, shape = setting)) + geom_point()

# truth
paste0(res_path, 'data/model/rho.csv') %>% read.csv(header = FALSE)
G = paste0(res_path, 'data/test_set/G.csv') %>% read.csv(header = FALSE)
Y = paste0(res_path, 'data/test_set/Y.csv') %>% read.csv(header = FALSE)
G$Category = factor(Y[[1]])
G %>% pivot_longer(!Category, names_to = "topic", values_to = "membership") %>%
  ggplot(aes_string(x="topic", y = "membership", color = "Category")) +
      geom_boxplot( ) + ggtitle("true G")
    
for(k0 in c(2, 5, 10)){
  p = paste0(res_path, 'supervised_', k0, '_', k0, '/G.csv') %>% 
    read.csv(header = FALSE) %>%
    mutate(Category = factor(Y[[1]])) %>%
    pivot_longer(!Category, names_to = "topic", values_to = "membership") %>%
    ggplot(aes_string(x="topic", y = "membership", color = "Category")) +
    geom_boxplot( ) + 
    ggtitle(paste0("G when k0 = ", k0))
  print(p)
}

for(k0 in c(2, 5, 10)){
  p = paste0(res_path, 'unsupervised_', k0, '_', k0, '/U.csv') %>% 
    read.csv(header = FALSE) %>%
    mutate(Category = factor(Y[[1]])) %>%
    pivot_longer(!Category, names_to = "topic", values_to = "membership") %>%
    ggplot(aes_string(x="topic", y = "membership", color = "Category")) +
    geom_boxplot( ) + 
    ggtitle(paste0("U when k0 = ", k0))
  print(p)
}


```



```{r}
# set 10:
# k0 = k1 = 2, d = 20, nlabel = 6
set_index = 10
set_type = 'unbalanced'
res_path = paste0('../res/', set_type, '_', set_index, '/')

accuracy %>% filter(set_index == 10 & set_type == 'balanced') %>% select(k0, k1, setting, method, accu) %>% ggplot(aes(x = k0, y = accu, color = method, shape = setting)) + geom_point()

# truth
paste0(res_path, 'data/model/rho.csv') %>% read.csv(header = FALSE)
G = paste0(res_path, 'data/test_set/G.csv') %>% read.csv(header = FALSE)
Y = paste0(res_path, 'data/test_set/Y.csv') %>% read.csv(header = FALSE)
G$Category = factor(Y[[1]])
G %>% pivot_longer(!Category, names_to = "topic", values_to = "membership") %>%
  ggplot(aes_string(x="topic", y = "membership", color = "Category")) +
      geom_boxplot( ) + ggtitle("true G")
    
for(k0 in c(2, 5, 10)){
  p = paste0(res_path, 'supervised_', k0, '_', k0, '/G.csv') %>% 
    read.csv(header = FALSE) %>%
    mutate(Category = factor(Y[[1]])) %>%
    pivot_longer(!Category, names_to = "topic", values_to = "membership") %>%
    ggplot(aes_string(x="topic", y = "membership", color = "Category")) +
    geom_boxplot( ) + 
    ggtitle(paste0("G when k0 = ", k0))
  print(p)
}

for(k0 in c(2, 5, 10)){
  p = paste0(res_path, 'unsupervised_', k0, '_', k0, '/U.csv') %>% 
    read.csv(header = FALSE) %>%
    mutate(Category = factor(Y[[1]])) %>%
    pivot_longer(!Category, names_to = "topic", values_to = "membership") %>%
    ggplot(aes_string(x="topic", y = "membership", color = "Category")) +
    geom_boxplot( ) + 
    ggtitle(paste0("U when k0 = ", k0))
  print(p)
}


```


It seems to agree with the conclusion that supervised model is more robust to misspecified k0 and k1.



```{r}
# set 45:
# k0 = k1 = 10, d = 20, nlabel = 6
set_index = 45
set_type = 'balanced'
res_path = paste0('../res/', set_type, '_', set_index, '/')

accuracy %>% filter(set_index == 10 & set_type == 'balanced') %>% select(k0, k1, setting, method, accu) %>% ggplot(aes(x = k0, y = accu, color = method, shape = setting)) + geom_point()

# truth
paste0(res_path, 'data/model/rho.csv') %>% read.csv(header = FALSE)
G = paste0(res_path, 'data/test_set/G.csv') %>% read.csv(header = FALSE)
Y = paste0(res_path, 'data/test_set/Y.csv') %>% read.csv(header = FALSE)
G$Category = factor(Y[[1]])
G %>% pivot_longer(!Category, names_to = "topic", values_to = "membership") %>%
  ggplot(aes_string(x="topic", y = "membership", color = "Category")) +
      geom_boxplot( ) + ggtitle("true G")

paste0(res_path, 'true/G_', set_index, '.csv') %>% 
    read.csv(header = FALSE) %>%
    mutate(Category = factor(Y[[1]])) %>%
    pivot_longer(!Category, names_to = "topic", values_to = "membership") %>%
    ggplot(aes_string(x="topic", y = "membership", color = "Category")) +
    geom_boxplot( ) + 
    ggtitle('G with true topics')

tem = paste0(res_path, 'true/U_', set_index, '.csv') %>% 
    read.csv(header = FALSE) 
tem[301:400,] %>%
    mutate(Category = factor(Y[[1]])) %>%
    pivot_longer(!Category, names_to = "topic", values_to = "membership") %>%
    ggplot(aes_string(x="topic", y = "membership", color = "Category")) +
    geom_boxplot( ) + 
    ggtitle('U with true topics')
    
for(k0 in c(2, 5, 10)){
  p = paste0(res_path, 'supervised_', k0, '_', k0, '/G.csv') %>% 
    read.csv(header = FALSE) %>%
    mutate(Category = factor(Y[[1]])) %>%
    pivot_longer(!Category, names_to = "topic", values_to = "membership") %>%
    ggplot(aes_string(x="topic", y = "membership", color = "Category")) +
    geom_boxplot( ) + 
    ggtitle(paste0("G when k0 = ", k0))
  print(p)
}

for(k0 in c(2, 5, 10)){
  p = paste0(res_path, 'unsupervised_', k0, '_', k0, '/U.csv') %>% 
    read.csv(header = FALSE) %>%
    mutate(Category = factor(Y[[1]])) %>%
    pivot_longer(!Category, names_to = "topic", values_to = "membership") %>%
    ggplot(aes_string(x="topic", y = "membership", color = "Category")) +
    geom_boxplot( ) + 
    ggtitle(paste0("U when k0 = ", k0))
  print(p)
}


```



```{r}
# set 45:
# k0 = k1 = 10, d = 20, nlabel = 6
set_index = 45
set_type = 'unbalanced'
res_path = paste0('../res/', set_type, '_', set_index, '/')

accuracy %>% filter(set_index == 10 & set_type == 'balanced') %>% select(k0, k1, setting, method, accu) %>% ggplot(aes(x = k0, y = accu, color = method, shape = setting)) + geom_point()

# truth
paste0(res_path, 'data/model/rho.csv') %>% read.csv(header = FALSE)
G = paste0(res_path, 'data/test_set/G.csv') %>% read.csv(header = FALSE)
Y = paste0(res_path, 'data/test_set/Y.csv') %>% read.csv(header = FALSE)
G$Category = factor(Y[[1]])
G %>% pivot_longer(!Category, names_to = "topic", values_to = "membership") %>%
  ggplot(aes_string(x="topic", y = "membership", color = "Category")) +
      geom_boxplot( ) + ggtitle("true G")
paste0(res_path, 'true/G_', set_index, '.csv') %>% 
    read.csv(header = FALSE) %>%
    mutate(Category = factor(Y[[1]])) %>%
    pivot_longer(!Category, names_to = "topic", values_to = "membership") %>%
    ggplot(aes_string(x="topic", y = "membership", color = "Category")) +
    geom_boxplot( ) + 
    ggtitle('G with true topics')

tem = paste0(res_path, 'true/U_', set_index, '.csv') %>% 
    read.csv(header = FALSE) 
tem[301:400,] %>%
    mutate(Category = factor(Y[[1]])) %>%
    pivot_longer(!Category, names_to = "topic", values_to = "membership") %>%
    ggplot(aes_string(x="topic", y = "membership", color = "Category")) +
    geom_boxplot( ) + 
    ggtitle('U with true topics')

for(k0 in c(2, 5, 10)){
  p = paste0(res_path, 'supervised_', k0, '_', k0, '/G.csv') %>% 
    read.csv(header = FALSE) %>%
    mutate(Category = factor(Y[[1]])) %>%
    pivot_longer(!Category, names_to = "topic", values_to = "membership") %>%
    ggplot(aes_string(x="topic", y = "membership", color = "Category")) +
    geom_boxplot( ) + 
    ggtitle(paste0("G when k0 = ", k0))
  print(p)
}

for(k0 in c(2, 5, 10)){
  p = paste0(res_path, 'unsupervised_', k0, '_', k0, '/U.csv') %>% 
    read.csv(header = FALSE) %>%
    mutate(Category = factor(Y[[1]])) %>%
    pivot_longer(!Category, names_to = "topic", values_to = "membership") %>%
    ggplot(aes_string(x="topic", y = "membership", color = "Category")) +
    geom_boxplot( ) + 
    ggtitle(paste0("U when k0 = ", k0))
  print(p)
}


```

Observation:

 - When k0 and k1 is large, it's hard for unsupervised model to learn meaningful topics and memberships, even if the true k0 and k1 is also large and even if the true topics are given. But unsupervised model can still capture some patterns of the data.

